Step 13: Forest Plots - Marginal Effects and PS Methods Comparison

### 2025-09-08 reanalysis-13-forest_plots.qmd
reorganize

  ame_plot <- ggplot(data_subgroups, aes(x = AME, y = Labels)) +
    geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
    geom_point(size = 3, color = "#2563eb") +
    geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "#2563eb") +
    geom_text(
      aes(x = Lower, label = sprintf("%.3f", Lower)), 
      hjust = 1.2, size = 3, color = "black"
    ) +
    geom_text(
      aes(x = Upper, label = sprintf("%.3f", Upper)), 
      hjust = -0.2, size = 3, color = "black"
    ) +
    geom_text(
      aes(x = AME, label = sprintf("%.3f", AME)), 
      vjust = -1, size = 3.5, fontface = "bold", color = "#2563eb"
    ) +
    scale_x_continuous(
      limits = c(min(data_subgroups$Lower) - 0.02, max(data_subgroups$Upper) + 0.02),
      breaks = seq(-0.1, 0.2, by = 0.05)
    ) +
    labs(
      title = "Average Marginal Effects of Lapsing Across Subgroups",
      subtitle = "Twang GBM with IPW - Pooled across 50 imputations",
      x = "Average Marginal Effect (AME)",
      y = "",
      caption = "AME > 0 indicates increased probability of vision impairment with lapsing"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 11, face = "bold"), 
      axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 10)),
      plot.subtitle = element_text(size = 11, hjust = 0.5, margin = margin(b = 20)),
      plot.margin = margin(20, 20, 20, 20),
      panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
      panel.grid.minor = element_blank(),
      plot.caption = element_text(size = 9, face = "italic", margin = margin(t = 20))
    )

@@

    ame_plot_with_overall <- ggplot(forest_data, aes(x = AME, y = Labels)) +
      geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
      geom_hline(yintercept = length(subgroup_order) + 0.5, 
                 color = "gray50", linetype = "dashed", alpha = 0.5) +
      geom_point(aes(size = point_size, color = point_color)) +
      geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = point_color), height = 0.2) +
      geom_text(
        aes(x = AME, label = sprintf("%.3f", AME), color = point_color), 
        vjust = -1, size = 3.5, fontface = "bold"
      ) +
      scale_color_identity() +
      scale_size_identity() +
      scale_x_continuous(
        limits = c(min(forest_data$Lower) - 0.02, max(forest_data$Upper) + 0.02),
        breaks = seq(-0.1, 0.2, by = 0.05)
      ) +
      labs(
        title = "Average Marginal Effects of Lapsing",
        subtitle = "Overall and Subgroup-Specific Effects",
        x = "Average Marginal Effect (AME)",
        y = "",
        caption = "Red: Overall effect | Blue: Subgroup effects"
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 11, face = "bold"), 
        axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 10)),
        plot.subtitle = element_text(size = 11, hjust = 0.5, margin = margin(b = 20)),
        plot.margin = margin(20, 20, 20, 20),
        panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
        panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, face = "italic", margin = margin(t = 20))
      )

@@

  ps_plot <- ggplot(ps_forest_data, aes(x = OR, y = label)) +
    geom_vline(xintercept = 1, color = "#be123c", linetype = "solid", size = 0.5) +
    geom_point(aes(size = size, color = color)) +
    geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = color), height = 0.2) +
    geom_text(
      aes(x = OR, label = sprintf("%.2f", OR), color = color), 
      vjust = -1, size = 3.5, fontface = "bold"
    ) +
    scale_color_identity() +
    scale_size_identity() +
    scale_x_continuous(
      trans = "log10",
      breaks = c(0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5),
      labels = c("0.8", "0.9", "1.0", "1.1", "1.2", "1.3", "1.4", "1.5")
    ) +
    labs(
      title = "Comparison of Propensity Score Methods",
      subtitle = "Odds Ratios for Vision Impairment (Lapse vs No Lapse)",
      x = "Odds Ratio (log scale)",
      y = "",
      caption = "OR > 1 indicates increased odds of vision impairment with lapsing\nRed: Primary method (Twang GBM) | Blue: Alternative methods"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 10), 
      axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 10)),
      plot.subtitle = element_text(size = 11, hjust = 0.5, margin = margin(b = 20)),
      plot.margin = margin(20, 20, 20, 20),
      panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
      panel.grid.minor = element_blank(),
      plot.caption = element_text(size = 9, face = "italic", margin = margin(t = 20))
    )

>>>>

  ame_plot <- ggplot(data_subgroups, aes(x = AME, y = Labels)) +
    geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
    geom_point(size = 3, color = "#2563eb") +
    geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "#2563eb") +
    geom_text(
      aes(x = Lower, label = sprintf("%.3f", Lower)), 
      hjust = 1.2, size = 3, color = "black"
    ) +
    geom_text(
      aes(x = Upper, label = sprintf("%.3f", Upper)), 
      hjust = -0.2, size = 3, color = "black"
    ) +
    geom_text(
      aes(x = AME, label = sprintf("%.3f", AME)), 
      vjust = -1, size = 3.5, fontface = "bold", color = "#2563eb"
    ) +
    scale_x_continuous(
      limits = c(min(data_subgroups$Lower) - 0.02, max(data_subgroups$Upper) + 0.02),
      breaks = seq(-0.1, 0.2, by = 0.05)
    ) +
    labs(
      title = "Average Marginal Effects of Lapsing Across Subgroups",
      subtitle = "Twang GBM with IPW - Pooled across 50 imputations",
      x = "Average Marginal Effect (AME)",
      y = "",
      caption = "AME > 0 indicates increased probability of vision impairment with lapsing"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 11, face = "bold"), 
      axis.title.x = element_text(size = 12, face = "bold", margin = ggplot2::margin(t = 20)),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = ggplot2::margin(b = 10)),
      plot.subtitle = element_text(size = 11, hjust = 0.5, margin = ggplot2::margin(b = 20)),
      plot.margin = ggplot2::margin(20, 20, 20, 20),
      panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
      panel.grid.minor = element_blank(),
      plot.caption = element_text(size = 9, face = "italic", margin = ggplot2::margin(t = 20))
    )

@@

    ame_plot_with_overall <- ggplot(forest_data, aes(x = AME, y = Labels)) +
      geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
      geom_hline(yintercept = length(subgroup_order) + 0.5, 
                 color = "gray50", linetype = "dashed", alpha = 0.5) +
      geom_point(aes(size = point_size, color = point_color)) +
      geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = point_color), height = 0.2) +
      geom_text(
        aes(x = AME, label = sprintf("%.3f", AME), color = point_color), 
        vjust = -1, size = 3.5, fontface = "bold"
      ) +
      scale_color_identity() +
      scale_size_identity() +
      scale_x_continuous(
        limits = c(min(forest_data$Lower) - 0.02, max(forest_data$Upper) + 0.02),
        breaks = seq(-0.1, 0.2, by = 0.05)
      ) +
      labs(
        title = "Average Marginal Effects of Lapsing",
        subtitle = "Overall and Subgroup-Specific Effects",
        x = "Average Marginal Effect (AME)",
        y = "",
        caption = "Red: Overall effect | Blue: Subgroup effects"
      ) +
      theme_minimal() +
      theme(
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 11, face = "bold"), 
        axis.title.x = element_text(size = 12, face = "bold", margin = ggplot2::margin(t = 20)),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = ggplot2::margin(b = 10)),
        plot.subtitle = element_text(size = 11, hjust = 0.5, margin = ggplot2::margin(b = 20)),
        plot.margin = ggplot2::margin(20, 20, 20, 20),
        panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
        panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, face = "italic", margin = ggplot2::margin(t = 20))
      )

@@

  ps_plot <- ggplot(ps_forest_data, aes(x = OR, y = label)) +
    geom_vline(xintercept = 1, color = "#be123c", linetype = "solid", size = 0.5) +
    geom_point(aes(size = size, color = color)) +
    geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = color), height = 0.2) +
    geom_text(
      aes(x = OR, label = sprintf("%.2f", OR), color = color), 
      vjust = -1, size = 3.5, fontface = "bold"
    ) +
    scale_color_identity() +
    scale_size_identity() +
    scale_x_continuous(
      trans = "log10",
      breaks = c(0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5),
      labels = c("0.8", "0.9", "1.0", "1.1", "1.2", "1.3", "1.4", "1.5")
    ) +
    labs(
      title = "Comparison of Propensity Score Methods",
      subtitle = "Odds Ratios for Vision Impairment (Lapse vs No Lapse)",
      x = "Odds Ratio (log scale)",
      y = "",
      caption = "OR > 1 indicates increased odds of vision impairment with lapsing\nRed: Primary method (Twang GBM) | Blue: Alternative methods"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 10),
      axis.text.y = element_text(size = 10), 
      axis.title.x = element_text(size = 12, face = "bold", margin = ggplot2::margin(t = 20)),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = ggplot2::margin(b = 10)),
      plot.subtitle = element_text(size = 11, hjust = 0.5, margin = ggplot2::margin(b = 20)),
      plot.margin = ggplot2::margin(20, 20, 20, 20),
      panel.grid.major.y = element_line(color = "gray90", linetype = "dotted"),
      panel.grid.minor = element_blank(),
      plot.caption = element_text(size = 9, face = "italic", margin = ggplot2::margin(t = 20))
    )
